---
title: "Introduction to Biostatistics - Learn Statistics the Right Way!"
subtitle: "Descriptive Statistics"
author: Justin Belair | Statistical Consultant | Author of Causal Inference in Statistics
output:
  beamer_presentation:
    theme: "metropolis"
    colortheme: "default"
    fonttheme: "structurebold"
    highlight: "tango"
    incremental: true
    fig_width: 7
    fig_height: 5
    fig_caption: true
    df_print: "kable"
    latex_engine: "xelatex"
    #keep_tex: true
    slide_level: 2
aspectratio: "169"
classoption: "professionalfonts"
header-includes:
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, pdfnewwindow=true}
  - \definecolor{dark_blue}{RGB}{29,42,59}
  - \definecolor{light_blue}{RGB}{227,237,255}
  - \setbeamercolor{frametitle}{bg=dark_blue, fg=light_blue}
  - \setbeamercolor{title}{fg=dark_blue}
  - \setbeamercolor{structure}{fg=dark_blue}
  - \usepackage{tikz}
  - \addtobeamertemplate{footline}{}{
      \begin{tikzpicture}[remember picture,overlay]
      \node[anchor=south west,yshift=-22pt,xshift=-5pt] at (current page.south west) {
        \href{https://justinbelair.ca/?utm_source=intro_biostats&utm_medium=slides&utm_campaign=logo_click}{\includegraphics[height=2.5cm]{C:/Users/justi/Documents/JBConsulting/Fiverr/Logo/JB_Statistical_Consulting_no_calligraphy_black.png}}
      };
      \end{tikzpicture}
    }
---

## Describing Distribution With Summary Statistics - Percentiles and Quantiles

- In `R`, the `quantile()` function is used to compute percentiles

```{r}
salary <- c(1000, 2000, 3000, 4000, 5000)
quantile(salary, probs = c(0, 0.01, 0.02, 0.03, 0.25, 0.5, 0.95))
```

```{python}
import numpy as np

salary = [1000, 2000, 3000, 4000, 5000]
probs = [0, 0.01, 0.02, 0.03, 0.25, 0.5, 0.95]
quantiles = np.percentile(salary, [p * 100 for p in probs])
print(quantiles)
```

- Let's look at the documentation by running `?quantile`

## Describing Distribution With Summary Statistics - Percentiles and Quantiles

- Let's try other types of quantiles

```{r}
salary <- c(1000, 2000, 3000, 4000, 5000)
quantile(salary, probs = c(0, 0.01, 0.02, 0.03, 0.25, 0.5, 0.95), type =  6)

quantile(salary, probs = c(0, 0.01, 0.02, 0.03, 0.25, 0.5, 0.95), type =  8)
```

```{python}
import numpy as np

salary = [1000, 2000, 3000, 4000, 5000]
probs = [0, 0.01, 0.02, 0.03, 0.25, 0.5, 0.95]

# R's type=6 corresponds to numpy's method='inverted_cdf'
quantiles_type6 = np.percentile(salary, [p * 100 for p in probs], method='inverted_cdf')
print(quantiles_type6)

# R's type=8 corresponds to numpy's method='median_unbiased' (default)
quantiles_type8 = np.percentile(salary, [p * 100 for p in probs], method='median_unbiased')
print(quantiles_type8)
```

## Contact Information and other clickable stuff

\small
- [**My Personal Website**](https://justinbelair.ca?utm_source=intro_biostats&utm_medium=slides_percentiles_quantiles_coding&utm_campaign=contact_info)
- If you are interested in Consulting, Training, Workshops, Public Speaking, Guest Lectures, and Other Custom Services in statistics and causal inference, please do not hesitate to contact me **belairjustin@gmail.com**
- The first chapter of my book, *Causal Inference in Statistics*, is available for free on my website by [**clicking here**](https://justinbelair.ca/causal-inference-in-statistics-book/?utm_source=intro_biostats&utm_medium=slides_percentiles_quantiles_coding&utm_campaign=contact_info)
- I share a montly newsletter to stats nerds, causal inference enthusiasts, and data scientists, [**subscribe here**](https://causal-inference-in-statistics.beehiiv.com/subscribe?utm_source=intro_biostats&utm_medium=slides_percentiles_quantiles_coding&utm_campaign=contact_info)
- I founded [**biostatistics.ca**](https://www.biostatistics.ca/?utm_source=intro_biostats&utm_medium=slides_percentiles_quantiles_coding&utm_campaign=contact_info), a blog and biostatistics community.
- I share content daily on [**Linkedin**](https://www.linkedin.com/in/justinbelair/)
- I created a [**Linkedin Causal Inference Group**](https://www.linkedin.com/groups/13065272/) with [**Aleksander** Molak](https://www.linkedin.com/in/aleksandermolak/)
- I created a [**Causal Inference Discord Community**](https://discord.com/invite/2xKvHMhYDg) which is free to join
