---
title: "Introduction to Biostatistics - Learn Statistics the Right Way!"
subtitle: "Descriptive Statistics"
author: Justin Belair | Statistical Consultant | Author of Causal Inference in Statistics
output:
  beamer_presentation:
    theme: "metropolis"
    colortheme: "default"
    fonttheme: "structurebold"
    highlight: "tango"
    incremental: true
    fig_width: 7
    fig_height: 5
    fig_caption: true
    df_print: "kable"
    latex_engine: "xelatex"
    #keep_tex: true
    slide_level: 2
aspectratio: "169"
classoption: "professionalfonts"
header-includes:
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, pdfnewwindow=true}
  - \definecolor{dark_blue}{RGB}{29,42,59}
  - \definecolor{light_blue}{RGB}{227,237,255}
  - \setbeamercolor{frametitle}{bg=dark_blue, fg=light_blue}
  - \setbeamercolor{title}{fg=dark_blue}
  - \setbeamercolor{structure}{fg=dark_blue}
  - \usepackage{tikz}
  - \addtobeamertemplate{footline}{}{
      \begin{tikzpicture}[remember picture,overlay]
      \node[anchor=south west,yshift=-22pt,xshift=-5pt] at (current page.south west) {
        \href{https://justinbelair.ca/?utm_source=intro_biostats&utm_medium=slides&utm_campaign=logo_click}{\includegraphics[height=2.5cm]{C:/Users/justi/Documents/JBConsulting/Fiverr/Logo/JB_Statistical_Consulting_no_calligraphy_black.png}}
      };
      \end{tikzpicture}
    }
---

```{r, echo=FALSE}
options(max.print = 100)
```

# What is "Typical" About a Given Distribution?    
## Describing Distributions With Summary Statistics - The Mean
- The most basic summary statistic is the *mean*  
  - It only makes sense to compute a mean for **quantitative variables**  

- The mean is the **average** of the data points  
  - It is calculated by summing all the values and dividing by the number of data points  

- Notation for the mean  
  - For a variable \( X \), we usually denote its (sample) mean by \( \bar{X} \) (read "x bar")  
  - Mathematically, the formula is  
    $$\bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i = \frac{x_1 + x_2 + \dots + x_n}{n}$$  
    where \( x_i \) is the \( i^{th} \) data point and \( n \) is the number of data points

## Describing Distributions With Summary Statistics - The Mean 
### Numerical Example

 - We roll a dice 5 times and note down the result 
 - This generates the following data set: $$2, 6, 1, 4, 5.$$ 
 - The (sample) mean of these 5 values is calculated as

## Describing Distributions With Summary Statistics - The Mean 

Let's perform the same computation with `R`.
```{r}
x <- c(2, 6, 1, 4, 5)
x_bar <- mean(x)
x_bar
```

## Describing Distributions With Summary Statistics - The Mean 

Let's perform the same computation with Python.
```{python}
import numpy as np

x = [2, 6, 1, 4, 5]
x_bar = np.mean(x)
print(x_bar)
```

## Describing Distributions With Summary Statistics - The Mean
### No Measure is Perfect
- The mean is called a *measure of central tendency*, since it seeks to describe the "center" of the distribution, what a "typical" value "looks like" 
- Consider the following example
  - We measure the weekly salary of 5 individuals in dollars : 1000, 2000, 1500, 1300, 50000 and calculate their mean

## In R
```{r}
salary <- c(1000, 2000, 1500, 1300, 50000)
salary_mean <- mean(salary)
salary_mean
```

## In Python
```{python}
import numpy as np

salary = [1000, 2000, 1500, 1300, 50000]
salary_mean = np.mean(salary)
print(salary_mean)
```

## Describing Distributions With Summary Statistics - The Mean
### No Measure is Perfect
- The mean is 11160 dollars--not quite representative of a "typical" salary right? 
- This is due to a single *outlier* in the data
  - E.g. we measured a very rich CEO who makes 50000 dollars a week, while the others have decent working-class salaries
  - The mean is heavily influenced by this single value
- This is *one* of the reasons I suggest complementing summary statistics with visual representations of data (we will get to this soon).

## Contact Information and other clickable stuff

\small
- [**My Personal Website**](https://justinbelair.ca?utm_source=intro_biostats&utm_medium=slides_mean&utm_campaign=contact_info)
- If you are interested in Consulting, Training, Workshops, Public Speaking, Guest Lectures, and Other Custom Services in statistics and causal inference, please do not hesitate to contact me **belairjustin@gmail.com**
- The first chapter of my book, *Causal Inference in Statistics*, is available for free on my website by [**clicking here**](https://justinbelair.ca/causal-inference-in-statistics-book/?utm_source=intro_biostats&utm_medium=slides_mean&utm_campaign=contact_info)
- I share a montly newsletter to stats nerds, causal inference enthusiasts, and data scientists, [**subscribe here**](https://causal-inference-in-statistics.beehiiv.com/subscribe?utm_source=intro_biostats&utm_medium=slides_mean&utm_campaign=contact_info)
- I founded [**biostatistics.ca**](https://www.biostatistics.ca/?utm_source=intro_biostats&utm_medium=slides_mean&utm_campaign=contact_info), a blog and biostatistics community.
- I share content daily on [**Linkedin**](https://www.linkedin.com/in/justinbelair/)
- I created a [**Linkedin Causal Inference Group**](https://www.linkedin.com/groups/13065272/) with [**Aleksander** Molak](https://www.linkedin.com/in/aleksandermolak/)
- I created a [**Causal Inference Discord Community**](https://discord.com/invite/2xKvHMhYDg) which is free to join