---
title: "Introduction to Biostatistics - Learn Statistics the Right Way!"
subtitle: "Modelling"
author: Justin Belair | Statistical Consultant | Author of Causal Inference in Statistics
output:
  beamer_presentation:
    theme: "metropolis"
    colortheme: "default"
    fonttheme: "structurebold"
    highlight: "tango"
    #incremental: true
    fig_width: 7
    fig_height: 3
    fig_caption: true
    df_print: "kable"
    latex_engine: "xelatex"
    #keep_tex: true
    slide_level: 2
aspectratio: "169"
classoption: "professionalfonts"
header-includes:
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, pdfnewwindow=true}
  - \definecolor{dark_blue}{RGB}{29,42,59}
  - \definecolor{light_blue}{RGB}{227,237,255}
  - \setbeamercolor{frametitle}{bg=dark_blue, fg=light_blue}
  - \setbeamercolor{title}{fg=dark_blue}
  - \setbeamercolor{structure}{fg=dark_blue}
  - \usepackage{tikz}
  - \addtobeamertemplate{footline}{}{
      \begin{tikzpicture}[remember picture,overlay]
        \node[anchor=south west,yshift=-22pt,xshift=-5pt] at (current page.south west) {
          \includegraphics[height=2.5cm]{C:/Users/justi/Documents/JBConsulting/Fiverr/Logo/JB_Statistical_Consulting_no_calligraphy_black.png}
        };
      \end{tikzpicture}
    }
---

```{r, echo=FALSE, warning = FALSE, message = FALSE}
options(max.print = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)

set.seed(1)
```

## Covariance and Correlation in the Diabetes Dataset
- In a sample, we use the sample covariance and sample correlation to estimate the population covariance and correlation
- The **sample covariance** is defined as:
$$s_{XY} = \frac{1}{n-1} \sum_{i=1}^{n} (X_i - \bar{X})(Y_i - \bar{Y})$$
- The **sample correlation** is defined as:
$$r_{XY} = \frac{s_{X, Y}}{s_X s_Y} = \frac{\sum_{i=1}^{n} (X_i - \bar{X})(Y_i - \bar{Y})}{(n-1)s_X s_Y}$$

# Example with a Dataset
## The Diabetes Dataset
- We will study a dataset that contains information about patients with diabetes, including their age, cholesterol levels, and other health metrics
- It is freely available at the [**Vanderbilt Biostatistics Datasets**](https://hbiostat.org/data/)
- It was studied in 
  - Willems JP, Saunders JT, DE Hunt, JB Schorling: Prevalence of coronary heart disease risk factors among rural blacks: A community-based study. Southern Medical Journal 90:814-820; 1997
  - Schorling JB, Roach J, Siegel M, Baturka N, Hunt DE, Guterbock TM, Stewart HL: A trial of church-based smoking cessation interventions for rural African Americans. Preventive Medicine 26:92-101; 1997.
- A [**data dictionary**](https://hbiostat.org/data/repo/cdiabetes) is available

## Visualizing the Diabetes Dataset
```{r, echo = FALSE}
data <- read.csv("diabetes.csv")
```

- A **scatter plot** is a type of plot that shows the relationship between two continuous variables
```{r, echo = FALSE}
chol.age.plot <- data %>%
  ggplot(aes(x = age, y = chol)) +
  geom_point() +
  labs(title = "Scatter Plot of Cholesterol vs Age",
       x = "Age (years)",
       y = "Cholesterol (mg/dL)") +
  theme_minimal()

chol.age.plot
```

## Covariance and Correlation in Sample Data - R

```{r, echo = FALSE}
data.complete <- data[complete.cases(data$chol) & complete.cases(data$age),]
```

\small
```{r}
var(data.complete$chol, data.complete$age)
cor(data.complete$chol, data.complete$age)

var(data.complete$chol, data.complete$age)/ 
  (sd(data.complete$chol) * sd(data.complete$age))
```

## Covariance and Correlation in Sample Data - Python
```{python,echo=FALSE}
import pandas as pd
data = pd.read_csv("diabetes.csv")
data_complete = data[['chol', 'age']].dropna()
```

\small
```{python}
covariance = data_complete['chol'].cov(data_complete['age'])
correlation = data_complete['chol'].corr(data_complete['age'])
print("Covariance:", covariance)
print("Correlation:", correlation)

calculated_correlation = covariance / (data_complete['chol'].std() * data_complete['age'].std())
print("Calculated Correlation:", calculated_correlation)
```
## Contact Information and other clickable stuff

\small
- [**My Personal Website**](https://justinbelair.ca?utm_source=intro_biostats&utm_medium=slides_covariance_correlation_data&utm_campaign=contact_info)
- If you are interested in Consulting, Training, Workshops, Public Speaking, Guest Lectures, and Other Custom Services in statistics and causal inference, please do not hesitate to contact me **belairjustin@gmail.com**
- The first chapter of my book, *Causal Inference in Statistics*, is available for free on my website by [**clicking here**](https://justinbelair.ca/causal-inference-in-statistics-book/?utm_source=intro_biostats&utm_medium=slides_covariance_correlation_data&utm_campaign=contact_info)
- I share a monthly newsletter to stats nerds, causal inference enthusiasts, and data scientists, [**subscribe here**](https://causal-inference-in-statistics.beehiiv.com/subscribe?utm_source=intro_biostats&utm_medium=slides_covariance_correlation_data&utm_campaign=contact_info)
- I founded [**biostatistics.ca**](https://www.biostatistics.ca/?utm_source=intro_biostats&utm_medium=slides_covariance_correlation_data&utm_campaign=contact_info), a blog and biostatistics community.
- I share content daily on [**Linkedin**](https://www.linkedin.com/in/justinbelair/)
- I created a [**Linkedin Causal Inference Group**](https://www.linkedin.com/groups/13065272/) with [**Aleksander Molak**](https://www.linkedin.com/in/aleksandermolak/)
- I created a [**Causal Inference Discord Community**](https://discord.com/invite/2xKvHMhYDg) which is free to join
- If you would just like to chat about science, stats, causal inference, or any other topic, please send me a DM on Linkedin or reach out via **belairjustin@gmail.com**.