---
title: "Introduction to Biostatistics - Learn Statistics the Right Way!"
subtitle: "Modelling"
author: Justin Belair | Statistical Consultant | Author of Causal Inference in Statistics
output:
  beamer_presentation:
    theme: "metropolis"
    colortheme: "default"
    fonttheme: "structurebold"
    highlight: "tango"
    #incremental: true
    fig_width: 7
    fig_height: 3
    fig_caption: true
    df_print: "kable"
    latex_engine: "xelatex"
    #keep_tex: true
    slide_level: 2
aspectratio: "169"
classoption: "professionalfonts"
header-includes:
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, pdfnewwindow=true}
  - \definecolor{dark_blue}{RGB}{29,42,59}
  - \definecolor{light_blue}{RGB}{227,237,255}
  - \setbeamercolor{frametitle}{bg=dark_blue, fg=light_blue}
  - \setbeamercolor{title}{fg=dark_blue}
  - \setbeamercolor{structure}{fg=dark_blue}
  - \usepackage{tikz}
  - \addtobeamertemplate{footline}{}{
      \begin{tikzpicture}[remember picture,overlay]
        \node[anchor=south west,yshift=-22pt,xshift=-5pt] at (current page.south west) {
          \includegraphics[height=2.5cm]{C:/Users/justi/Documents/JBConsulting/Fiverr/Logo/JB_Statistical_Consulting_no_calligraphy_black.png}
        };
      \end{tikzpicture}
    }
---

```{r, echo=FALSE, warning = FALSE, message = FALSE}
options(max.print = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(stringr)

set.seed(1)
```

## Logistic Regression - continued
```{r, echo = FALSE}
data.prostate <- read.csv("prostate.csv") %>%
  mutate(death = str_trim(str_extract(status, "^[^-]*")),
         death = ifelse(death == "dead", 1, 0))
```
- We will study the prostate data set from the Vanderbilt Biostatistics Library
- It was studied in Byar DP, Green SB (1980): Bulletin Cancer, Paris 67:477-488
- More information can be found [**here**](https://hbiostat.org/data/repo/prostate) and the data dictionary can be found [**here**](https://hbiostat.org/data/repo/cprostate)
- I converted it to .csv before importing it to R, and cleaned up the `status` variable to obtain a binary variable `death` indicating whether the patient died or not

## Logistic Regression - continued
- We will model death as a function of age, weight, size of tumor, cancer stage, and patient fitness
```{r}
glm.logistic <- glm(death ~ age + wt + stage + rx, 
                    data = data.prostate, family = binomial(link = "logit"))
```
## Logistic Regression - continued
\tiny
```{r, echo = FALSE}
summary(glm.logistic)
```

## Logistic Regression
- The coefficients are interpreted as the change in the *log-odds* of the outcome for a one-unit change in the predictor, holding all other predictors constant
- We can also exponentiate the coefficients to obtain the *odds ratios*, which are easier to interpret
\small
```{r, echo = FALSE}
exp(coef(glm.logistic)[-1]) %>% #we remove the intercept
  round(3)
```
## Logistic Regression - continued
- We cannot perform an ANOVA per se, since the model is not a linear model
- We can use the `car` package to obtain an ANOVA-like table, based on an **analysis-of-deviance**, a slightly more complicated procedure that we will not discuss in detail
```{r}
car::Anova(glm.logistic)
```

## Logistic Regression - continued
- Here we see that the main-effect of medication is very high and significant
- Be careful with causal interpretations of this model

## Contact Information and other clickable stuff

\small
- [**My Personal Website**](https://justinbelair.com?utm_source=intro_biostats&utm_medium=slides_logistic_regression&utm_campaign=contact_info)
- If you are interested in Consulting, Training, Workshops, Public Speaking, Guest Lectures, and Other Custom Services in statistics and causal inference, please do not hesitate to contact me **belairjustin@gmail.com**
- The first chapter of my book, *Causal Inference in Statistics*, is available for free on my website by [**clicking here**](https://justinbelair.ca/causal-inference-in-statistics-book/?utm_source=intro_biostats&utm_medium=slides_logistic_regression&utm_campaign=contact_info)
- I share a monthly newsletter to stats nerds, causal inference enthusiasts, and data scientists, [**subscribe here**](https://causal-inference-in-statistics.beehiiv.com/subscribe?utm_source=intro_biostats&utm_medium=slides_logistic_regression&utm_campaign=contact_info)
- I founded [**biostatistics.ca**](https://www.biostatistics.ca/?utm_source=intro_biostats&utm_medium=slides_logistic_regression&utm_campaign=contact_info), a blog and biostatistics community.
- I share content daily on [**Linkedin**](https://www.linkedin.com/in/justinbelair/)
- I created a [**Linkedin Causal Inference Group**](https://www.linkedin.com/groups/13065272/) with [**Aleksander Molak**](https://www.linkedin.com/in/aleksandermolak/)
- I created a [**Causal Inference Discord Community**](https://discord.com/invite/2xKvHMhYDg) which is free to join
- If you would just like to chat about science, stats, causal inference, or any other topic, please send me a DM on Linkedin or reach out via **belairjustin@gmail.com**.